A multidimensional approach for measuring meritocratic beliefs: Advantages, limitations and alternatives to the ISSP social inequality survey

### 2023-10-06 paper-blinded-RR-IJS.Rmd
ajustes pre render

The item used by @mijs_paradox_2019 is part of an item's battery which is available in several international surveys (as ISSP), usually called “reasons to get ahead”. This battery displays a series of indicators related to what people consider important to get ahead: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender. Therefore, considering only one of the items means that other aspects associated with talent, such as education, would not be deemed meritocratic. As he points out: "Hard work is arguably the most meritocratic part of Michael Young's equation: 'Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5). In such conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Young's' original conceptualization). This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only (or main) dimension behind the concept of meritocracy?

@@

```{r scat1, fig.cap="Cross-national comparison of non-meritocratic and meritocratic perceptions with meritocratic preference",fig.height=6, fig.width=12}
  load(file = "input/data/proc/data_s1.RData")
  if (!require("pacman")) install.packages("pacman")
#  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
    "html" #if html set "html" in format
  } else if(knitr::is_latex_output()) {
    "latex"#if latex set "latex" in format
  }

@@

>>>>

The item used by @mijs_paradox_2019 is part of an item's battery which is available in several international surveys (as ISSP), usually called “reasons to get ahead”. In the ISSP, this battery belongs to the social inequality module---see @roberts_issp_2023 for a recent overview of this module. This battery displays a series of indicators related to what people consider important to get ahead: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender. Therefore, considering only one of the items means that other aspects associated with talent, such as education, would not be deemed meritocratic. As he points out: "Hard work is arguably the most meritocratic part of Michael Young's equation: 'Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5). In such conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Young's' original conceptualization). This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only (or main) dimension behind the concept of meritocracy?

@@

```{r scat1, fig.cap="Cross-national comparison of non-meritocratic and meritocratic perceptions with meritocratic preference",fig.height=6, fig.width=12}
  load(file = "input/data/proc/data_s1.RData")
  if (!require("pacman")) install.packages("pacman")
 pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,gridExtra)
  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
    "html" #if html set "html" in format
  } else if(knitr::is_latex_output()) {
    "latex"#if latex set "latex" in format
  }

@@

```{r fit-country-s2, eval=FALSE, include=FALSE}
if (!require("pacman")) install.packages("pacman")
  #---Load packages and data -------------------------------------#
  pacman::p_load(dplyr,lavaan,semTools,knitr,kableExtra,stats,gtools,haven)
  options(knitr.kable.NA = '')
  load(file = "input/data/proc/data_s1.RData")

# meassurement model: Second order
  cfa_perpref2 <- '
  	  # latent variables
  	  percmerit =~ hwork + ambition
  	  parents =~ wealthy + pareduc
  	  backgrd =~ race + gender
  	  networks=~ people + polcone
  	  percnmerit=~ parents + backgrd + networks
  	  prefmerit =~ welljob + hardjob
  	  prefnmerit=~ family + child
  	  '

  data_s1$country <- countrycode::countrycode(sourcevar = data_s1$v5,origin = "iso3n",destination = "country.name")
# estimar modelo por paises
df1<-data_s1 %>% select(country,wealthy:hardjob)

# sjmisc::frq(df1$country)
# 
# Argentina      <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Argentina",],ordered = T)
# Australia      <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Australia",],ordered = T)
# Austria        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Austria",],ordered = T)
# Belgium        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Belgium",],ordered = T)
# # Bulgaria       <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Bulgaria",],ordered = T)
# Chile          <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Chile",],ordered = T)
# China          <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="China",],ordered = T)
# Croatia        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Croatia",],ordered = T)
# Cyprus         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Cyprus",],ordered = T)
# Czechia        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Czechia",],ordered = T)
# Denmark        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Denmark",],ordered = T)
# Estonia        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Estonia",],ordered = T)
# Finland        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Finland",],ordered = T)
# France         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="France",],ordered = T)
# Germany        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Germany",],ordered = T)
# Hungary        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Hungary",],ordered = T)
# # Iceland        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Iceland",],ordered = T)
# Israel         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Israel",],ordered = T)
# # Italy          <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Italy",],ordered = T)
# # Japan          <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Japan",],ordered = T)
# Latvia         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Latvia",],ordered = T)
# # Lithuania      <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Lithuania",],ordered = T)
# NewZealand    <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="New Zealand",],ordered = T)
# Norway         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Norway",],ordered = T)
# Philippines    <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Philippines",],ordered = T)
# Poland         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Poland",],ordered = T)
# # Portugal       <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Portugal",],ordered = T)
# Russia         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Russia",],ordered = T)
# Slovakia       <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Slovakia",],ordered = T)
# Slovenia       <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Slovenia",],ordered = T)
# SouthAfrica   <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="South Africa",],ordered = T)
# SouthKorea    <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="South Korea",],ordered = T)
# Spain          <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Spain",],ordered = T)
# # Sweden         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Sweden",],ordered = T)
# Switzerland    <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Switzerland",],ordered = T)
# Taiwan         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Taiwan",],ordered = T)
# Turkey         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Turkey",],ordered = T)
# Ukraine        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Ukraine",],ordered = T)
# UnitedKingdom <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="United Kingdom",],ordered = T)
# # UnitedStates  <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="United States",],ordered = T)
# # Venezuela      <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Venezuela",],ordered = T)
# # 
# # 
# # 
# lavaan::fitmeasures(Argentina     )
# lavaan::fitmeasures(Australia     )
# lavaan::fitmeasures(Austria       )
# lavaan::fitmeasures(Belgium       )
# # lavaan::fitmeasures(# Bulgaria    )
# lavaan::fitmeasures(Chile         )
# lavaan::fitmeasures(China         )
# lavaan::fitmeasures(Croatia       )
# lavaan::fitmeasures(Cyprus        )
# lavaan::fitmeasures(Czechia       )
# lavaan::fitmeasures(Denmark       )
# lavaan::fitmeasures(Estonia       )
# lavaan::fitmeasures(Finland       )
# lavaan::fitmeasures(France        )
# lavaan::fitmeasures(Germany       )
# lavaan::fitmeasures(Hungary       )
# # lavaan::fitmeasures(# Iceland     )
# # lavaan::fitmeasures(Israel        )
# # lavaan::fitmeasures(Italy         )
# # lavaan::fitmeasures(# Japan       )
# lavaan::fitmeasures(Latvia        )
# # lavaan::fitmeasures(# Lithuania   )
# lavaan::fitmeasures(NewZealand    )
# lavaan::fitmeasures(Norway        )
# lavaan::fitmeasures(Philippines   )
# lavaan::fitmeasures(Poland        )
# # lavaan::fitmeasures(# Portugal    )
# lavaan::fitmeasures(Russia        )
# lavaan::fitmeasures(Slovakia      )
# lavaan::fitmeasures(Slovenia      )
# lavaan::fitmeasures(SouthAfrica   )
# lavaan::fitmeasures(SouthKorea    )
# lavaan::fitmeasures(Spain         )
# # lavaan::fitmeasures(# Sweden      )
# lavaan::fitmeasures(Switzerland   )
# lavaan::fitmeasures(Taiwan        )
# lavaan::fitmeasures(Turkey        )
# lavaan::fitmeasures(Ukraine       )
# lavaan::fitmeasures(UnitedKingdom )
# # lavaan::fitmeasures(UnitedStates)
# # lavaan::fitmeasures(Venezuela     )

df2<-data_s1 %>% select(country,wealthy:hardjob) %>% 
  filter(!country %in% c("Bulgaria","Iceland","Israel","Italy","Japan",
                         "Lithuania","Portugal","Sweden","United States","Venezuela"))

df_paises <- split(df2, df2$country)
results = lapply(df_paises, function(df2) lavaan::cfa(cfa_perpref2,data = df2,ordered = T))

#extraer fitmeasures por paises
results_list<- lapply(results, function(results) data.frame(t(lavaan::fitmeasures(results)[c("chisq","df","cfi","tli","rmsea")])))
#extraer nobs por paises
nobs_list <- lapply(results, function(results) data.frame(obs=t(lavaan::nobs(results))))
nobs_n <- data.table::rbindlist(nobs_list) %>% 
  mutate(country=names(nobs_list)) %>%
  select(country,everything())

  # extract fit indices from models and add to table
  sum_fit <- 
  data.table::rbindlist(results_list) %>% 
  mutate(country=names(results_list)) %>% 
  select(country,everything())
  # Customize object
  sum_fit$nobs <- nobs_n$obs
  sum_fit <- dplyr::select(sum_fit,country,"chisq","df", nobs,"cfi","tli","rmsea")
  sum_fit <- sum_fit %>% dplyr::arrange(chisq)
  colnames <- c("País","$\\chi^2$","gl","$N$","CFI","TLI","RMSEA")
  # Create table
  tab_cfa_count_1d <-
    kable(
      sum_fit,
      # format=table_format,
      digits = 3,
      booktabs = T,
      col.names = colnames,
      caption = "Summary of country-specific fit for the ISSP",
      escape = FALSE
    ) %>% 
    kable_styling(full_width = F,
                  font_size = 10,
                  latex_options = "HOLD_position",
                  bootstrap_options=c("striped", "bordered"))
  tab_cfa_count_1d
```

```{r invariance-s2, eval=FALSE, include=FALSE}
df2<-data_s1 %>% select(country,wealthy:hardjob) %>% 
  filter(!country %in% c("Bulgaria","Iceland","Israel","Italy","Japan",
                         "Lithuania","Portugal","Sweden","United States","Venezuela","Finland"))

  inv01 <- semTools::measurementInvariance(model=cfa_perpref2,
                                           data=df2,
                                           group="country",estimator = "ML",strict=TRUE,quiet = T)

  var_or <- c("hwork","ambition","wealthy", "pareduc","race",
              "gender","people","polcone","welljob" ,"hardjob","family","child")

  inv02 <- semTools::measurementInvarianceCat(model=cfa_perpref2,
                                              data=df2,
                                           group="country",estimator = "DWLS",
                                           strict=TRUE,quiet = T,ordered=var_or)

  ### Check an alternative method
  configural        <- lavaan::cfa(cfa_perpref2, data=df2, group = "country", estimator="ML")
  weak.invariance   <- lavaan::cfa(cfa_perpref2, data=df2, group = "country",group.equal = "loadings", estimator="ML")
  strong.invariance <- lavaan::cfa(cfa_perpref2, data=df2, group = "country",group.equal = c( "loadings", "intercepts"), estimator="ML")
  strict.invariance <- lavaan::cfa(cfa_perpref2, data=df2, group = "country",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
